---
output:
  html_document: default
  pdf_document: default
---

```{r include=F}
    library(tidyverse)
    library(ggplot2)
```


# Arbeitsblatt 3

## Frequency Tables and Distributions

### Aufgabe 1

-   Consider the results of the national elections in Germany in 2013 and 2017:

    | Party     | Results 2013 | Results 2017 |
    |-----------|--------------|--------------|
    | CDU       | 26.8         | 34.1         |
    | SPD       | 20.5         | 25.7         |
    | AfD       | 12.6         | 4.7          |
    | FDP       | 10.7         | 4.8          |
    | DIE LINKE | 9.2          | 8.6          |
    | GRUENE    | 8.9          | 8.4          |
    | CSU       | 6.2          | 7.4          |
    | OTHER     | 5.0          | 6.2          |

    Summarize the results of 2017 in a pie and a bar chart. Compare the results in 2013 and 2017 with an appropriate bar chart.

```{r}
    Results <- tibble(Party=c("CDU", "SPD", "AfD", "FDP", "DIE LINKE",             "GRUENE", "CSU", "Others"), 
                  '2013'=c(26.8, 20.5, 12.6, 10.7, 9.2, 8.9, 6.2, 5.0), 
                  '2017'=c(34.1, 25.7, 4.7, 4.8, 8.6, 8.4, 7.4, 6.2))
```
```{r}
    Results <- Results %>% mutate(diff = Results$'2013' - Results$'2017')
    Results$'2017' %>% pie(labels=Results$Party)

    Results$'2017' %>% barplot(names.arg = Results$Party, ylab = "Voter Percent", ylim = c(0,40))

    Results$diff %>% barplot(names.arg=Results$Party,
                ylim=c(-10, 10),
                xlab="Parties", ylab="Difference to 2013")
```

### Aufgabe 2
The data shown in the list are the times in milliseconds it took one of
us to move the mouse over a small target in a series of 20 trials. The
times are sorted from shortest to longest.
568, 577, 577, 581, 640, 641, 645, 657, 673, 696, 703, 720, 728, 728, 728,
729, 777, 808, 824, 825, 865, 975, 1007, 1007

```{r}
cumfreqs <- tibble(times.mills= c(568, 577, 577, 581, 640, 641, 645, 657, 673, 696, 703, 720, 728, 728, 728, 729, 777, 808, 824, 825, 865, 975, 1007, 1007))
cumfreqs <- cumfreqs %>% count(times.mills) %>% 
mutate (
    abs.freq = n ,
    rel.freq = abs.freq / sum(abs.freq),
    cum.rel.freq = cumsum (rel.freq)
) %>% select(-n)
```
```{r echo=F}
    cumfreqs
```

- Compute and draw the cumulative frequency distribution.

```{r}
H <- ecdf(cumfreqs$times.mills)
```
```{r eval=F}
plot.ecdf(cumfreqs$times.mills ,
                    xlab = "time" , ylab = "H(x)" ,
                    main = "ungrouped Data")
```
```{r}
plot(H)
```


    
- Compute using the cumulative frequency distribution the proportion of response times
- less equal 800
```{r}
H(800)
```
        
- greater than 725
```{r}
1-H(725)
```
        
- greater than 642 and less equal 777
```{r}
H(777) - H(642)
```
        
- equal 696
```{r}
H(800) - H(696) + sum(cumfreqs$times.mills == 696) / length(cumfreqs$times.mills)
```
        

