---
output:
  html_document: default
  pdf_document: default
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Arbeitsblatt 2
## Tidy Data
### Aufgabe 1

```{r}
student1 <- tibble(
student = c("Adam","Bernd","Christian","Doris"),
algebra = c(NA, 5, 3, 4),
analysis = c(2, NA, 1,3),
diskrete.math = c(3,NA,2,4))

student2 <- tibble(
name = rep(c("Adam", "Bernd", "Christian", "Doris"), each = 2),
type = rep(c("height", "weight"), 4),
measure = c(1.83, 81, 1.75, 71, 1.69, 55, 1.57, 62))

student3 <- tibble(
name = c("Adam", "Bernd", "Christian", "Doris"),
ratio = c("81/1.83", "71/1.75", "55/1.69", "62/1.57"))
```

```{r echo=F}
    student1
    student2
    student3
```

- Describe for every dataset what the dataset contains? What are
the variables and what are the observations?

```
```

- Why are these datasets are not tidy?

```
```

- Make a tidy version of all datasets.

```{r}
student1 %>%
    gather(c("algebra", "analysis", "diskrete.math"), key = "subject", value = "grades", na.rm = T)

student1 %>% pivot_longer(cols = c(algebra, analysis, diskrete.math), names_to = "exam",values_to = "grade", values_drop_na = T)

student2 %>% 
    spread(key = "type", value= "measure")

student2 %>% 
    pivot_wider(names_from = "type", values_from = "measure")

student3 %>%
    separate(col = "ratio", into = c("weight", "height"), sep = "/") %>%
    mutate(ratio = as.double(weight) / as.double(height))
```

### Aufgabe 2

- Calculate the value of $sin(log(\sqrt{5 + 3}))$ directly and using the %>%-operator
```{r}
    5 %>% `+`(3) %>% sqrt %>% log %>% sin
```
- Define a vector v with values 0.5,1,1.5,...,5 and calculate the by 2
digits rounded sum of the logarithms of the squared values of v
with nested operations and using the %>%-operator.

```{r}
    v <- seq(0.5, 5, 0.5)
    round(sum(log10(v ** 2)), 2)
    (v ** 2) %>% log10 %>% sum %>% round(2)
    v %>% '^'(2) %>% log10 %>% sum %>% round(2)
```

### Aufgabe 3
Create a tibble df with the data of 10 students, i.e. with 10 rows and
the columns id (values 1,2,..., 10), sex (values are “f” and “m”, age
(integer values between 20 and 35) and score1 (integer values between
0 and 25). You can choose arbitrary values in the columns. If you do
not like coding the values by hand you can use:

```{r}
    df <- tibble(id = 1:10,
              sex = sample(x = c('m','f'), size = 10, replace = T),
              age = round(runif(10, 20, 35)),
              score1 = round(runif(10, 0, 25)))
```
```{r echo=F}
    df
```

- Select the date of all male students.

```{r}
    df %>% filter(sex == "m")
```

- Add the data of a new student with id = 11, sex = “m”, age =
25 and score1 = 4

```{r}
    df <- df %>% add_row(id = 11, sex = "m", age = 25, score1 = 4)
```
```{r echo=F}
    df
```

- Add two columns score2 and score3 with random integer numbers
between 0 and 25.

```{r}
    df <- df %>% add_column(score2 = sample(0:25, 11, T), score3 = sample(0:25, 11, T))
```
```{r echo=F}
    df
```

- Add a column containing sum of all scores.

```{r}
    df <- df %>% mutate(sum = rowSums(df[,4:6])) 
    df %>% mutate(sum = score1 + score2 + score3)
    df %>% rowwise() %>% mutate(sum = sum(score1, score2, score3))
```

- Add a column which denote the grades according to the scheme
$$
grad = \begin{cases}
    5  & if & score sum \leq 37\\
    4  & if & 37 < score sum \leq 45\\
    3  & if & 45 < score sum \leq 55\\
    2  & if & 55 < score sum \leq 65\\
    1  & if & score sum > 65
\end{cases}
$$

```{r}
    df <- df %>% mutate(grad = case_when(sum <= 37 ~ 5,
                                         sum > 37 & sum <=45 ~ 4,
                                         sum > 45 & sum <= 55 ~ 3,
                                         sum > 55 & sum <= 65 ~ 2,
                                         sum > 65 ~ 1))
```

- Find the values of the variables id, sex and grade sorted by the
values of sex of all students who have passed.

```{r}
    df %>% filter(grad <= 4) %>% select(id, sex, grad) %>% arrange(sex)
```

- Calculate the mean, minimum, maximum and median of the variable sum of scores grouped by the variable sex.

```{r}
    df$sum %>% aggregate( by=list(df$sex), FUN=\(x) {c(mean= mean(x), median= median(x), min= min(x), max= max(x))})

    df %>% group_by(sex) %>% summarise(mean=mean(sum), min=min(sum), max=max(sum), median=median(sum))
    
    df %>% group_by(sex) %>% reframe(mean=mean(sum), min=min(sum), max=max(sum), median=median(sum))

```

### Aufgabe 4

```{r}
no <- 30
exercise.results <- tibble(
    stud.id = 1:no,
    group = sample(x=c("A","B","C"), size=no, replace = TRUE),
    ex1 = sample(x=1:10, size=no, replace = TRUE),
    ex2= sample(x=1:10, size=no, replace = TRUE),
    ex3 = sample(x=1:10, size=no, replace = TRUE),
    ex4 = sample(x=1:10, size=no, replace = TRUE),
    ex5 = sample(x=1:10, size=no, replace = TRUE)
)
```

- Apply n() and count() to get the number of students in the different groups. What are the difference between n() and count()?

```{r}
exercise.results %>% count(group)
exercise.results %>% group_by(group) %>% summarize(n = n())
```


- Add the variables sum.scores and mean.scores containing the sum
and the of the scores in the exercises for every student by applying the the functions sum() and mean(). What is the result if
rowwise() is appplied before the mutate()?

```{r}
exercise.results %>% rowwise %>% mutate(sum.scores = sum(ex1:ex5), mean.scores = mean(ex1:ex5))
```


### Aufgabe 5
```{r include=F}
    library(nycflights13)
```

- Find all flights with more than 2 hours arrival delay.

```{r}
    nycflights13::flights %>% filter(arr_delay > 120)
```

- Find all flights with more than 2 hours arrival delay and no departure delay.

```{r}
    nycflights13::flights %>% filter(arr_delay > 120 & dep_delay <= 0)
```

- Find all flights from United, American and Delta with no arrival delay.

```{r}
    nycflights13::flights %>% filter(carrier %in% c('UA', 'AA', 'DL')) %>% filter(arr_delay <= 0)
```

- Find all flights from United, American and Delta in the month
May with more than 5 hours arrival delay sorted by carrier and
flight number.

```{r}
    nycflights13::flights %>% filter(carrier %in% c('UA', 'AA', 'DL')) %>% filter(arr_delay >= 5) %>% filter(month == 5) %>% arrange(carrier, flight)
```


- Exchange the values of departure time and arrival time in minute
after midnight.

```{r}
    conv_minutes <- \(x) {(x%%100) + (x%/%100) * 60}

    nycflights13::flights %>% mutate(arr_time = conv_minutes(arr_time),
                                     dep_time = conv_minutes(dep_time))
```


- Add a column speed which denotes the average speed of the flight
and determine the carrier, flight of the top 10 values of speed.

```{r}
    nycflights13::flights %>% mutate(speed = round(distance / (air_time / 60))) %>% arrange(desc(speed)) %>% select(carrier, flight, speed) %>% slice(0:10)
```

- Find a list of carriers with a column ratio which denotes the number of flights with arr delay less than 10 minutes to the total
number of flights. The list should be sorted by ratio.

```{r}
    nycflights13::flights %>% 
    filter(!is.na(arr_delay)) %>% 
    group_by(carrier) %>% 
    summarise(nof = n(), 
              ndel = sum(arr_delay < 10), 
              del_ratio = round(ndel/nof, 3)) %>% 
    arrange(desc(del_ratio)) 
```


- Find a list which denotes for every month the carrier with highest
ratio which denotes the number of flights with arr delay less than
10 minutes to the total number of flights. The list should have
the columns month, carrier, number of flights of the carrier in that
month and ratio.

```{r}
    nycflights13::flights %>%
    filter(!is.na(arr_delay)) %>% 
    group_by(carrier, month) %>% 
    summarize(nof = n(), 
              ndel = sum(arr_delay < 10), 
              del_ratio=round(ndel/nof, 3)) %>% 
    arrange(desc(del_ratio)) %>% 
    group_by(month) %>% 
    slice(which.max(del_ratio))
```

- Find a table with the number of cancelled flights (dep delay =
NA), the number of flights with no dep delay ( —dep delay—
≤ ± 5 minutes and the means of dep delay, arr delay per month
and day.

```{r}
    canceled_flights <- nycflights13::flights %>% 
    group_by(month, day) %>% 
    summarise(nof_canc = sum(is.na(dep_delay)), .groups = "keep")

    punctual_flights <- nycflights13::flights %>% 
    group_by(month, day) %>% 
    summarise(nof_canc = sum(is.na(dep_delay)), .groups = "keep")

    mean_delays <- nycflights13::flights %>%
    filter(!is.na(dep_delay) & !is.na(arr_delay)) %>% 
    group_by(month, day) %>% 
    summarise(mean_dep_delay = mean(dep_delay), mean_arr_delay = mean(arr_delay), .groups = "keep")
    
    merge(canceled_flights, punctual_flights) %>% 
    merge(mean_delays) %>%
    arrange(month, day)
```

- Determine a table that shows, for each airline (carrier), the flight
connection given by the airports of dest und origin that occurred
most frequently in 2013. The table should contain only the columns
names of airline, destination, origin and frequency and be sorted
by frequency in descending order. You can find the names of the
carrier from the data set airlines and the names of the airports
from the data set airports.

```{r}
    nycflights13::flights %>% 
    group_by(carrier, origin, dest) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(carrier) %>%
    filter(n == max(n)) %>% 
    ungroup() %>% 
    left_join(nycflights13::airlines, by= "carrier") %>% 
    rename(Airline = name) %>% 
    left_join(nycflights13::airports, by= c("origin"="faa")) %>% 
    rename(Airport.Origin = name) %>% 
    left_join(nycflights13::airports, by= c("dest"="faa")) %>% 
    rename(Airport.Destination = name) %>% 
    select(n, Airline, Airport.Origin, Airport.Destination) %>% 
    arrange(desc(n))
```

### Aufgabe 6

```{r include=F}
    library(tidyr)
```

- Consider the dataset who of the package tidyr.
country, iso2, iso3, and year are already variables, so they can
be left as is. But the columns from new sp m014 to newrel f65
encode four variables in their names:
    - The new /new prefix indicates these are counts of new cases. This dataset only contains new cases, so we’ll ignore it here because it’s constant.
    - sp/rel/ep describe how the case was diagnosed.
    - m/f gives the gender.
    - 014/1524/2535/3544/4554/65 supplies the age range.
Break these variables up by specifying multiple column names and
make the data set tidy.

```{r}
     tidyr::who %>% pivot_longer(cols = new_sp_m014:newrel_f65, 
                                   names_to = c("type", "sex", "age"),
                                   names_pattern = "new_?(.*)_(.)(.*)",
                                   values_to = "cases", values_drop_na = T)
```

- The anscombe dataset contains four pairs of variables (x1 and y1,
x2 and y2, etc.) that underlie Anscombe’s quartet, a collection of
four datasets that have the same summary statistics (mean, sd,
correlation etc), but have quite different data.
Produce a dataset with columns set, x and y.

```{r}
    anscombe %>% pivot_longer(
        cols=everything(),
        cols_vary = "slowest",
        names_to = c(".value", "set"),
        names_pattern = "(.)(.)"
    )
```


- Apply the following R statements. The data set production contains the combination of product,country, and year. Widen the data so we have one column for each combination of product and country.

```{r, label='production'}
    production <- expand_grid(
        product = c("A", "B"),
        country = c("AI", "EI"),
        year = 2000:2014) %>%
    filter((product == "A" & country == "AI") | product == "B") %>%
    mutate(production = rnorm(nrow(.)))
```

```{r}
    production %>% pivot_wider(values_from = production, names_from = c(product, country))
```

- The data set warpbreaks gives the number of warp breaks per
loom, where a loom corresponds to a fixed length of yarn for every
combination of wool (A and B) and tension (L, M, H).
Produce a data set with the columns tension, A, B with the means
of the breaks.

```{r}
    warpbreaks %>% pivot_wider(names_from = wool,
                               values_from = breaks,
                               values_fn = mean)
```

