---
title: "Sheet9"
output: html_document
---

```{r, include=F}
    knitr::opts_chunk$set(echo=T, error=T)
```

```{r, include=F}
    library(tidyverse)
```



## 1)
```{r}
    p.eval <- \(p.val, alpha){if(p.val >= alpha) " fail to" else ""}

    m.1 <- c(5.46, 5.34, 4.34, 4.82, 4.4, 5.12, 5.69, 5.53, 4.77, 5.82)
    m.2 <- c(5.45, 5.31, 4.11, 4.69, 4.18, 5.05, 5.72, 5.54, 4.62, 5.89, 5.6, 5.19, 3.31, 4.43, 5.3, 4.09)
    sig.1 <- 0.5
    sig.2 <- 0.6
    alpha <- 0.05
    r <- qnorm(alpha)
    
    t <- (mean(m.1) - mean(m.2)) / sqrt((sig.1 ^ 2 / length(m.1)) + (sig.2 ^ 2 / length(m.2)))
    
    p.value <- pnorm(t)
    
    
    t <= r
```


We`r p.eval(p.value, alpha)` reject $H_0$

## 2)
```{r}
    g.1 <- c(7.06, 11.84, 9.28, 7.92, 13.5, 3.98, 3.82, 7.34, 8.7, 9.24, 4.86, 3.32, 12.78, 12, 5.24, 11.4, 6.56, 9.04, 7.72, 9.26, 7.88, 8.6, 9.3, 8.42, 8.54)
    g.2 <- c(8.68, 6, 6.3, 10.24, 10.88, 5.36, 7.82, 4.7, 9.02, 9.78, 6.9, 5.8, 13.56, 10.32, 13.3, 11.38, 7.94, 10.74, 13.68, 14.92, 7.42, 10.36, 10.54, 5.22, 13.74, 12.98, 10.34, 10.02, 17.8, 13.04, 5.2, 9.4, 11.18, 12.68, 12.36)
    m <- 25
    n <- 35
    alpha <- 0.05
```

$H_0: \sigma(g.1) = \sigma(g.2)$

$H_1: \sigma(g.1) \neq \sigma(g.2)$

```{r}
var.test(g.1, g.2, alternative="two.sided", conf.level=1-alpha)$p.value < alpha
```
We fail to reject $H_0$


$H_0: g.1 \geq g.2$

$H_1: g.1 < g.2$

```{r}
t.test(g.1, g.2, alternative = "less", var.equal = T,conf.level = 1-alpha)$p.value < alpha
```
We reject $H_0$. Therefore method B is better

## 3)

```{r}
n <- 10
alpha <- 0.05
g.w <- c(16, 15, 11, 20, 19, 14, 13, 15, 14, 16)
g.a <- c(13, 13, 10, 18, 17, 11, 10, 15, 11, 16)

test.result <- t.test(g.w, g.a, alternative="two.sided", paired = T, var.equal = T, conf.level=1-alpha)
```

$H_0: \mu (g.w) = \mu (g.a)$

$H_1: \mu (g.w) \neq \mu (g.a)$
```{r}
test.result$p.value < alpha
```
We reject $H_0$. Therefore alcohol worsens the motor ability.

$t: `r test.result$statistic`$

$p: `r test.result$p.value`$

## 4)
```{r}
g.a <- c(102.4, 101.3, 97.6, 98.2, 102.3, 99.1, 97.8, 103.9, 101.6, 100.1)
g.b <- c(98.4, 101.7, 100.5, 99.3, 100.6, 99.6, 102.2, 101.1, 99.9, 101.0)
n <- length(g.a)
alpha <- 0.05
```

$H_0: \sigma(a) \leq \sigma(b)$

$H_1: \sigma(a) > \sigma(b)$

```{r}
test.result <- var.test(g.a, g.b, alternative="greater", conf.level = 1-alpha)$p.value < alpha
```
We reject $H_0$. Therefore scale B is statistically better.

## 5)

```{r}
g.1 <- c(7.2, 4.1, 5.5, 4.5, 5.7, 3.8, 4.6, 6.0, 5.2, 5.4)
g.2 <- c(5.3, 4.4, 5.0, 3.5, 3.9, 4.9, 5.6, 2.5, 4.0, 3.6)
alpha <- 0.1
```

### a.
$H_0: \sigma(a) = \sigma(b)$

$H_1: \sigma(a) \neq \sigma(b)$

```{r}
var.test(g.1, g.2, alternative="two.sided", conf.level = 1-alpha)$p.value < alpha
```
We fail to reject $H_0$. Therefore the dispersion is the same.

### b.

```{r}
alpha <- 0.025
```

$H_0: \mu(a) \leq \mu(b)$

$H_1: \mu(a) > \mu(b)$

```{r}
t.test(g.1, g.2, alternative="greater", var.equal = T, conv.level=1-alpha)$p.value < alpha
```
We reject $H_0$. Therefore the weight gain from composition A is greater than weight gain from composition B


## 14)

### a.
```{r}
data <- read_delim("magnets_pain.csv", delim = ";", col_types = "iiii")
```

### b.
```{r}
boxplot(data$Score_2 ~ data$Active)
```


```{r}
data %>% group_by(Active) %>% 
        summarise(mean = mean(Score_2),
                  min = min(Score_2),
                  max = max(Score_2),
                  sd = sd(Score_2),
                  q1 = quantile(Score_2, probs=0.25),
                  q2 = quantile(Score_2, probs=0.5),
                  q3 = quantile(Score_2, probs=0.75),
                  iqr = q3 - q1) -> summ
summ
```
### c.

$H_0: \sigma(G.1) = \sigma(G.2)$

$H_1: \sigma(G.1) < \sigma(G.2) \cap \sigma(G.1) > \sigma(G.2)$
```{r}
alpha <- 0.05
var.test(data$Score_2 ~ data$Active, alternative="two.sided", conv.level= 1-alpha)$p.value < alpha
```

We Reject $H_0$, therefore the $\sigma$ are not equal


$H_0: G.1 = G.2$

$H_1: G.1 < G.2 \cap G.1 > G.2$
```{r}
t.test(data$Score_2 ~ data$Active, alternative="two.sided", var.equal=F, conv.level = 1-alpha)
```


