---
title: "WS23 Probe"
author: 'Matrikelnr.: 1481589, Name: Mert Kaan Demirel'
date: "2025-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
```

```{r}
library(tidyverse)
library(TeachingDemos)
```

# Descriptive

# 1.

```{r}
data <- read.csv("corona.csv", sep = ",") %>% 
    as_tibble()
data
```

# 2.

continent: qualitative, nominal

location: qualitative, nominal

year: quantitative, interval

month: quantitative, interval

day: quantitative, interval

total_cases: quantitative, ratio

new_cases: quantitative, ratio

total_deaths: quantitative, ratio

new_deaths: quantitative, ratio

reproduction_rate: quantitative, ratio

population: quantitative, ratio

# 3.

```{r}
octger <- data %>% 
    filter(month == 10 & location == "Germany") %>% 
    select(day,new_deaths, new_cases)
octger
```

# 4.

```{r}
new <- data %>% 
    group_by(month, location) %>% 
    summarise(
        new_cases = sum(new_cases),
        new_deaths = sum(new_deaths)
    )
    
new
```

# 5.

```{r}
total <- data %>%
    filter(!is.na(new_cases) | !is.na(new_deaths)) %>% 
    group_by(month) %>% 
    summarise(
        minC = min(new_cases),
        minD = min(new_deaths),
        maxC = max(new_cases),
        maxD = max(new_deaths),
        meanC = mean(new_cases),
        meanD = mean(new_deaths),
        medianC = median(new_cases),
        medianD = median(new_deaths)
    )
total
```

# 6.

```{r}
seven <- data %>% 
    filter(location == "Germany" & year == 2020)    
seven
```
WTF is the true solution.

# 7.

```{r}
lm <- lm(seven$new_deaths ~ seven$new_cases)
lm
```

a.

```{r}
a <- lm$coefficients[1]
a
```

```{r}
b <- lm$coefficients[2]
b
```

b.

```{r}
plot(seven$new_cases, seven$new_deaths)
abline(a = a, b = b)
```

c.

```{r}
cor(lm$model) # Correlation

cor(lm$model)**2 # Determination
```

d.
```{r}
a + b * 20000
```

e.

Correlation: New Deaths correlate proportional to new cases.

Determination: Half of the deaths are explained by the cases.


# 8.

```{r}
march <- read.csv("corona_march_2020.csv", sep = ",") %>% 
    as_tibble()
march
```
Cases and deaths are in the same column. Days are split into separate columns.
```{r}
march.tidy <- march %>% 
    pivot_longer(cols = X1:X31, names_to = "day", names_prefix = "X", values_to = "temp", values_drop_na = T) %>% 
    separate(temp, into = c("new_cases", "new_deaths"), sep = "/", convert = T) %>% 
    mutate(
        day = as.integer(day)
    )
march.tidy
```

# Probability

# Ex. 1

a.

```{r}
1 - pbinom(q = 40, size = 100, prob = 0.5)
```

b.

```{r}
dmultinom(x = c(30, 50, 20), prob = c(0.4, 0.5, 0.1))
```

c.

```{r}
dgeom(x = 19, 0.1)
```

d.

```{r}
try <- tibble(
    n = 100:150,
    prob = pnorm(80, mean = n * 0.5, sd = sqrt(0.25 * n))
) %>% 
    filter(prob <= 0.9) %>% 
    head(1) %>% 
    select(n)
try
```

# Ex. 2

a.

```{r}
p <- c(0.2, 0.4, 0.25, 0.15)
n <- c(3, 4, 2, 2.5)

EX <- sum(p * n)
EX2 <- sum(p * n^2)

VarX <- sum(p * n^2) - sum(p * n)^2

mean100 <- 100 * EX
sd100 <- sqrt(100 * VarX)

1 - pnorm(300, mean = mean100, sd = sd100)
```

b.

```{r}
qnorm(0.95, mean = mean100, sd = sd100)
```

# Inferential

# Ex. 1

```{r}
sample <- c(0.92,0.83,0.81,0.70,0.88,0.73,1.05,0.91,0.83,0.67,0.94,0.90,0.91,
0.83,0.84,0.96,0.87,0.91,0.98,0.84,0.88,0.76,0.99,0.89,0.82)

sd <- 0.2
```

a.

```{r}
alpha <- 0.01

test <- z.test(x = sample, stdev = sd, alternative = "two.sided", conf.level = 1-alpha)
test$conf.int
```
Confidence Interval: [0.7629668, 0.9690332]

b.

```{r}
length <- 0.9690332 - 0.7629668
length
```

c.

```{r}
length <- 0.15
```

i.

```{r}
q <- qnorm(1 - (alpha / 2))
n <- ceiling(((2 * q * sd) / length)^2)
n
```

ii.

```{r}
n <- length(sample)
z_value <- (length * sqrt(n)) / (2 * sd)
alpha <- 2 * (1 - pnorm(z_value))
1-alpha
```
Confidence Level is 93.9%

d.

$H_0$: mu <= 0.9
$H_1$: mu > 0.9

```{r}
alpha <- 0.05

test <- t.test(x = sample, mu = 0.9, alternative = "greater", conf.level = 1-alpha)
t <- test$statistic
r <- test$conf.int[1]
t
t > r
test$p.value < alpha
```
As both tests are false, we don't reject $H_0$

# Ex. 2

```{r}
x <- c(98.32,97.26,99.85,99.52,95.73,95.56,100.49,98.19,95.16,98.26,
96.46,100.23,99.76,98.58,97.43)
```

a.

$H_0$: mu = 100

$H_1$: mu != 100

b.

t-test

c.

```{r}
alpha <- 0.05

test <- t.test(x = x, mu = 100, alternative = "two.sided", conf.level = 1-alpha)
test$statistic < test$conf.int[1] | test$statistic > test$conf.int[2]
```

```{r}
test$p.value
test$p.value < alpha
```
As both tests equal true, we reject $H_0$

d.

```{r}
y <- c(100.14,100.05,96.51,98.70,98.22,101.06,103.55,100.16,100.60,
102.85,103.15,100.66,102.52,102.09,100.84)
```

$H_0$: mu(x) >= mu(y)

$H_1$: mu(x) < mu(y)

e.

$H_0$: sigma(x) = sigma(y)

$H_1$: sigma(x) != sigma(y)

```{r}
alpha <- 0.1

test <- var.test(x = x, y = y, alternative = "two.sided", conf.level = 1-alpha)
test$statistic < test$conf.int[1] | test$statistic > test$conf.int[2]
test$p.value < alpha
```
As both are false, we don't reject $H_0$, therefore the variances are equal.

f.

```{r}
alpha <- 0.05

test <- t.test(x = x, y = y, alternative = "less", paired = F, var.equal = T, conf.level = 1-alpha)
test$statistic < test$conf.int[2]
test$p.value
test$p.value < alpha
```
As both are true, we reject $H_0$