---
title: "R  Cheatsheet"
output:
  html_document:
    df_print: paged
---

```{r include=F}
    knitr::opts_chunk$set(eval=FALSE)
```

# Descriptive

## Input from file

- Read data from csv
```{r}
    read.csv() #for Comma separated values
    read.csv2() #for Semicolon separated values
```

## Import Libraries

```{r}
    library()
```

## Create Table

```{r}
    t <- tibble(col1 = 1:29, ...)
```

## Create Vector
```{r}
    vec <- c(v1, v2, ...)
```

## Replicate Values
```{r}
    rep(c("A", "B"), each=2) # AABB
    rep(c("A", "B"), 2)      # ABAB
    rep(1, 5)                # 1 1 1 1 1
```

## Sequencing Values
```{r}
    seq(from = 1, to = 5, by = 0.5) #
    1:5
```
## Random Values

```{r}
    round(runif(10, 20, 35)) #10 Random floats between 20 and 35 rounded to int
    sample(x = 1:10, size = 5, replace=T) #5 values sampled from a range from 1 to 10, multiples allowed 
```

## Edit Table Columns
```{r}
    mutate(col1 = 4)
    rowwise() %>% mutate(col1 = sum(col2, col3)) # sum of col2 & col3 per row
```

## Add Row to Table
```{r}
    add_row(col1 = 34)
```

## Switch Case
```{r}
    case_when(col1 > 3 ~ "Greater 3", ...)
```

## if else
```{r}
    if_else(col1 >= 5, "Greater Equal 5", "Smaller 5")
```

## Filter data
```{r}
    filter(col1 > 6)
```

### Value is NA
```{r}
    is.na(col1) # true if value is NA
```

### Value in Vector
```{r}
    col1 %in% c("A", "B")
```

## Viewing specific Columns
```{r}
    select(col1, col2)
```

## Viewing specific Columns
```{r}
    slice()
```


## Sorting
```{r}
    arrange(col1)
```

## Grouping data
```{r}
    group_by(col1)
```

## Summarise
```{r}
    summarise(col1 = mean(col2))
```

## Counting Data
```{r}
    summarise(col1 = n()) # counts items in group
    count(group1) 
```


## Split Column into multiple
```{r}
    separate(col1, into=c("new_col2", "new_col3"), sep="/")
```

## Combine multiple Columns to one

```{r}
    unite(col1, old_col2, old_col3, sep="/")
```

## Expanding Tables 
```{r}
    pivot_wider(names_from=col1, values_from = col2)
```
    
names_from: column name where the new column names should be drawn from
values_from: column name where values should be drawn from

```{r}
    pivot_longer(cols = col1:col3, names_to = new_col4, names_pattern = "(.)(.)", values_to = new_col5)
```
    
cols: vector of colums where values are drawn from
names_to: vector of new column names for naming colums
names_pattern: regex how to split the column names from <cols>
values_to: column name for value column

## contengency/indifference tables, chi-sqared test
```{r}
    test_result <- chisq.test(<colums_separated_by_comma>)
    test_result$observed # contengency table, use addmargin for margins
    test_result$expected # indifference table, use addmargin for margins 
    test_result$statistic # chi-squared value
```

## Linear Regression 
$y = a + b * x$
```{r}
    lin_reg <- lm(col2 ~ col1)
    a <- lin_reg$coefficients[1] # intercept
    b <- lin_reg$coefficients[2] # slope of regression line
```

## Covariance
```{r}
    cov(x, y)
```

## Corelation Coefficient
```{r}
    cor(x ,y)
```

## Determination Coefficient
```{r}
    cor(x, y) ^ 2
```



# Inferential Statistics
## Hypotheses Test

$H_0: P \leq 80\%$ gegen hypothese

$H_1: P > 80\%$ hypothese

$n=200, \overline{x} = 172, \alpha = 0.05$

```{r}
    alpha <- 0.05 #significance level
    qnorm(1-alpha) # bound 1-tailed
    qnorm(1-alpha/2) #upper bound 2-tailed
    qnorm(alpha/2) #lower bound 2-tailed
```

$\frac{r_n - p_0}{\sqrt{\frac{p_0 * (1-p_0)}{n}}}$

$\frac{}{}$
