---
title: "Exam24"
output: 
    html_document:
        theme:
            version: 4
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=T)
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

## Aufgabe 1

The file europe.data.csv is a subset from World Bank, population
19602020, containing only european countries. The data set has 61 rows
and 47 columns.

-   year: years from 1960 to 2020
-   Albania: population of Albania in year
-   Andorra: population of Andorra in year
-   ...
-   United Kingdom: population of United Kingdom in year

### a)

Specify type (qualitativ, quantitave) and scale (nominal, ordinal,
interval, ratio) of all variables.

| Column       | Type         | Scale   |
|--------------|--------------|---------|
| Year         | qualitative  | nominal |
| Country_Name | quantitative | ratio   |

### b)

Import europe.data.csv into a tibble.

```{r}
    europe.data <- read.csv("europe.data.csv", sep = ";", dec=",")
```

### c)

The data set is messy. Why? Make the data tidy and change the values of
year and the population value to integer.

The Columns with Country Names represent more of a Observation

```{r}
    europe.data %>% 
        mutate(year = str_remove(year, "year_") %>% as.integer(year), 
               United.Kingdom = as.integer(United.Kingdom)) %>% 
        pivot_longer(Albania:United.Kingdom, names_to = "country", values_to = "population") -> tidy
    tidy
```

### d)

Create a table containing the size of the population and the gain from
2000 to 2020 of the top 5 countries with biggest population in 2020. The
table should be sorted according the population gain.

```{r}
    tidy %>% 
        filter(year == 2020) %>% 
        top_n(5, population) %>% 
        select(country) %>% 
        as_vector() -> top5

    tidy %>% 
        filter(country %in% top5 & year %in% c(2000,2020)) %>% 
        group_by(country) %>% 
        mutate(growth = population[2] - population[1]) %>%
        filter(year == 2020) %>% 
        select(-year) %>% 
        arrange(-growth)
```

### e

Consider only the german population and draw a scatterplot with the
years as x values and the population values as y-values and add the
regression line.

```{r}
    tidy %>% 
        filter(country == "Germany") %>% 
        select(!country) -> g.pop
    
    g.pop %>% 
        plot()
    lm(population ~ year, data=g.pop) -> linear_model
    abline(linear_model, col="red")
```

### f)

Determine the coefficients of the linear regression
$population = a + b * year$ and interpret the values.

```{r}
    linear_model$coefficients
```

### g)

Determine the coefficient of determination and interpret the value.

```{r}
    cor(g.pop$year, g.pop$population) ^ 2
```

### h)

Make a prediction of the population in 2021 based on the linear
regression.

```{r}
    a <- linear_model$coefficients[1] 
    b <- linear_model$coefficients[2] 
    a + b * 2021
```

## Aufgabe 2

At a university, 30% of all students belong to faculty A, 25% of all
students belong to faculty B, 30% of all students belong to faculty C
and 15% of all students belong to faculty D. 100 students from the
university are selected at random.

```{r}
    p.a <- 0.3
    p.b <- 0.25
    p.c <- 0.3
    p.d <- 0.15
    n <- 100
```

### a)

What is the probability that:

#### i)

between 25 and 35 of the selected students belong to faculty A.

```{r}
    pbinom(35, n, p.a) - pbinom(24, n, p.a)
```

#### ii)

35 students of the selected students belong to faculty A, 30 students to
faculty B, 20 students to Faculty C and 15 students to Faculty D.

```{r}
    dmultinom(c(35, 30, 20, 15), prob = c(p.a, p.b, p.c, p.d))
```

#### iii)

if the students are selected randomly one after the other, the 5th
student is the first student from faculty C or D.

```{r}
    (1-(p.c+p.d))^4*(p.c+p.d)
```

### b)

The randomly selected students from faculties B and C are all to be
tested together in a room with 80. What is the maximum sample size such
that the room is sufficient with a probability of 0.5?

```{r}
n <- 1:2000
tibble(
    n = n,
    v =  pbinom(80, n, prob = 0.5)
) %>% filter(v < 0.9) %>% head(1)
```

## Aufgabe 3

At a speed check, 60% of all vehicles comply with the speed limit of 50
km/h. 20% of all vehicles have a speed between 50 km/h and 60 km/h, 15%
of all vehicles have a speed between 60 km/h and 70 km/h and 5% have a
speed above 70 km/h. A penalty of 20 euros is imposed for exceeding the
speed limit by a maximum of 10 km/h. The amount of the penalty is 50
euros if the speed limit is exceeded by between 10 km/h and 20 km/h and
100 euros if the speed limit is exceeded by more than 20 km/h.

```{r}
    penalties <- tibble(prob = c(0.6, 0.2, 0.15, 0.05), 
                       fine=c(0, 20, 50, 100))
    n <- 500
```

### a)

f 500 vehicles are measured, what is the approximate distribution of the
penalties imposed?

```{r}
    E.penalty <- (penalties$fine * penalties$prob) %>% sum()
    E.penalty.2 <- (penalties$fine^2 * penalties$prob) %>% sum()
    Var.penalty <- E.penalty.2 - E.penalty^2
    E.penalty; Var.penalty
```

### b)

What is the approximate probability that, for 500 vehicles measured, the
amount of the penalties imposed will be between 8000 and 9000 Euro.

```{r}
    pnorm(9000, mean = n*E.penalty, sd = sqrt(n*Var.penalty)) -
      pnorm(8000, mean = n*E.penalty, sd = sqrt(n*Var.penalty))
```

### c)

What is the minimum amount of the penalty imposed for 500 vehicles
measured with a probability of 0.9?

```{r}
    qnorm(0.9, mean = n*E.penalty, sd = sqrt(n*Var.penalty))
```

Assume that the speeds of the vehicles are independent, identically
distributed random variables.

## Aufgabe 4

The file runs.csv contains the results of a run with 186 participants.
In file the times, the age, the gender and the names of all participants
are given.

### a)

Import runs.csv into a tibble.

```{r}
    read.csv2("runs.csv") -> runs
```

### b)

Assume that the variable time is approximately normally distributed.
Evaluate a 95% confidence interval for the unknown expected value of the
times of the females.

```{r}
    runs %>% filter(gender == "F") %>% select(time) %>% as_vector() -> time.female
    t.test(time.female, alternative = "two.sided", conf.level = 1-0.05)$conf.int
```

### c)

Give an interpretation of the confidence interval.

We are 95% confident that the population parameter is between 103.2382
and 109.5149.

### d)

Change the confidence level so that the length of the confidence
interval is 4.

```{r}
    sd.time <- sd(time.female)
    mean.time <- mean(time.female)
    nobs <- length(time.female)
    conf.level <- 2 * pt(2*nobs^0.5/sd.time, df = nobs-1)-1
    conf.level
```

### e)

One year ago the mean of the time of females was 101.98. Conduct an
appropriate statistical test on a 5 percent level to check whether the
expected value is bigger than last year. Specify the null hypothesis.
What is the p-value? What is your decision?

```{r}
    p <- t.test(time.female, alternative = "greater", mu=101.98)$p.value
    p
    p <= 0.05
```

Reject $H_0$

## Aufgabe 5

### a)

Import the file mpg.csv into a tibble.

```{r}
    mpg <- read.csv2("mpg.csv")
```

### b)

Interpret the side-by-side boxplot.

Cars with manual Transmission are more efficient per gallon.

### c)

Assume that the random variable highway miles per gallon (hwy) is
normally distributed. Use an appopriate statistical test to check
whether the variance can be assumed to be equal or not on a 10% level.
Specify the null hypothesis. What is the p-value? What is your decision?

```{r}
    var.test(mpg$hwy ~ mpg$trans, alternative= "two.sided", conf.level=0.90)
```

### d)

You want to check the conjecture that cars with automatic transmission
will have a higher consumption. Specify an appropriate statistical test
to check the conjecture by the given date and conduct the test on a 1%
level. Specify the null hypothesis. What is the p-value? What is your
decision?

```{r}
    t.test(mpg$hwy ~ mpg$trans, alternative= "less", conf.level=0.99)
```
